---
output:
  html_document: default
  pdf_document: default
---
Red Wine Data Exploration by Elias Paz
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(ggcorrplot)
library(GGally)
library(pastecs)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
df <- read.csv('wineQualityReds.csv')

```



# Introduction:
I will be exploring the Red Wine Quality data set. This data contains data from 11 different variables that help determine the quality of a red wine.  

# Questions
In exploring this data set, i will seek to answer the following questions: \
1. What does the Quality of Wine distribution look like? \
2. What properties and their respective values make up a good wine? \
2. What properties and their respective values make up a defective wine? \
\

# Univariate Plots Section \

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
# Exploration of data structure and variable summaries
str(df)
```

This data set has 1599 observations with 14 variables. 

```{r echo=FALSE} 
# Exploration of variable summaries
summary(df$quality)
summary(df$alcohol)
```

```{r echo=FALSE}
bargraph.fun <- function(var1,df){
  p1 <- ggplot(aes(x=var1),data=df) +
    geom_bar(stat = "count", color = 'blue', fill = 'black') 
  return (p1)
  
}

```


```{r echo=FALSE}
histogram.fun <- function(var1,df){
  
  p1 <- ggplot(aes(x = var1), data = df) +
    geom_histogram(color = 'black', fill = 'green') 
  return (p1)
}
```

```{r echo=FALSE}
desc_stats.fun <- function(var1){
  options(scipen=100)
  options(digits=2)
  return(stat.desc(var1))
  }
```





```{r echo=FALSE}
# Plot of Quality ratings
bargraph.fun(df$quality,df) +
  geom_vline(xintercept = mean(df$quality), linetype=1, color='red', alpha=1) +
  geom_vline(xintercept = median(df$quality), linetype=2, color='red', alpha=1) +
  coord_cartesian(xlim=c(2,9)) 

table(df$quality)
quality.labs <- paste('low', 'medium','high')
df$qaulity.bucket <- cut(df$quality, breaks = c("2","4","6","8"), labels = c("low","medium","high"))
table(df$qaulity.bucket)
summary(df$quality)

df.s1 <- subset(df, quality < mean(quality))
df.s2 <- subset(df, quality > mean(quality))

length(df.s1$quality)
length(df.s2$quality)
```

After running some summaries on the data, the first categorical variable I wanted to explore was the Quality. \ 
From the graph above, it looks like majority of wines are above the mean of 5.636 which is displayed in the solid red line.\
The median is the stripped line.  I wonder how what the alcohol percentage is for the sampled wine. 

```{r echo=FALSE}
histogram.fun(df$alcohol,df) +
    geom_vline(xintercept = mean(df$alcohol), linetype=1, color='red', alpha=1) +
  scale_y_log10()
```




From the graph above, it looks like most wines samples had an alcohol percentage of about 9-10.5%. \
There are a few outliers at above 14% I wonder what other variables such as pH and the different acids look like. 

```{r echo=FALSE}
# pH plot
ggplot(aes(x = pH), data = df) +
  geom_freqpoly()
```
As stated in the data set explanation, most wines have a pH between 3 and 4 and that is evident in the graph above. \
It is said that pH represents how much acid is in a wine regardless of taste. \
Next lets see the total acidity, tritratable acid, of the sampled wines. 

```{r echo=FALSE}
#creating new variable. Total Acidity. 
df$total.acidity <- df$fixed.acidity + df$volatile.acidity + df$citric.acid
ggplot(data = df, aes(x = total.acidity)) +
  geom_freqpoly()
summary(df$total.acidity)
```

As seen above, majority of wines have 10.07 g/mL of total acidity. It is said that tritation measures how strong the acid tastes.\


```{r echo=FALSE}
p1 <- histogram.fun(df$fixed.acidity,df) +
  coord_cartesian(xlim = c(4,quantile(df$fixed.acidity,.99))) +
  xlab("Fixed Acidity")

p2 <- histogram.fun(df$volatile.acidity,df) +
  coord_cartesian(xlim = c(0.1,quantile(df$volatile.acidity,.99))) +
  xlab("Volatile Acidity")
  
p3 <- histogram.fun(df$citric.acid,df) +
  coord_cartesian(xlim = c(0,quantile(df$citric.acid,.99))) +
  xlab("Citric Acid")
p4 <- histogram.fun(df$total.acidity,df) +
  coord_cartesian(xlim = c(4,quantile(df$total.acidity,.99))) +
  xlab ("Total Acidity")

grid.arrange(p1,p2,p3,p4, ncol=2)

```



The following observations from the plots above are: \
1. Fixed acidity peaks between 4 and 9 g/mL \
2. Volatile acidity peaks between .2 and .7 g/mL \
3. Majority of wines have a citric acid level of .25 g/mL or less \
4. There is signifcanty more fixed acidity in the sampled wines than volatile or citric acids. \
5. Total Acidity seems to have a normal distribtuion. \
6. There appears to be outliers present across all Acidity categories. \
\

Lets see some descriptive statistics about the different acidity levels \




```{r echo=FALSE}
# descriptive stats for fixed.acidity
desc_stats.fun(df$fixed.acidity)

#subsetting data to count how many entries are above and below the mean and median
df.s7.mean <- subset(df, fixed.acidity > mean(fixed.acidity))
df.s8.mean <- subset(df, fixed.acidity < mean(fixed.acidity))
df.s7.median <- subset(df, fixed.acidity > median(fixed.acidity))
df.s8.median <- subset(df, fixed.acidity < median(fixed.acidity))
length(df.s7.mean$fixed.acidity)
length(df.s8.mean$fixed.acidity)
length(df.s7.median$fixed.acidity)
length(df.s8.median$fixed.acidity)
```
From the table above, it appears there are outliers in the data as the graph suggests. The max value of 15.9 is more than 3 STDs from the mean

```{r echo=FALSE}
# descriptive stats for volatile.acidity
desc_stats.fun(df$volatile.acidity)

#subsetting data to count how many entries are above and below the mean and median

df.s9.mean <- subset(df, volatile.acidity > mean(volatile.acidity))
df.s10.mean <- subset(df, volatile.acidity < mean(volatile.acidity))
df.s9.median <- subset(df, volatile.acidity > median(volatile.acidity))
df.s10.median <- subset(df, volatile.acidity < median(volatile.acidity))
length(df.s9.mean$volatile.acidity)
length(df.s10.mean$volatile.acidity)
length(df.s9.median$volatile.acidity)
length(df.s10.median$volatile.acidity)
```
From the table above, it appears there are outliers in the data as the graph suggests. The max value of 1.58 is more than 3 STDs from the mean

```{r echo=FALSE}
# descriptive stats for citric.acid
desc_stats.fun(df$citric.acid)

#subsetting data to count how many entries are above and below the mean and median

df.s11.mean <- subset(df, citric.acid > mean(citric.acid))
df.s12.mean <- subset(df, citric.acid < mean(citric.acid))
df.s11.median <- subset(df, citric.acid > median(citric.acid))
df.s12.median <- subset(df, citric.acid < median(citric.acid))
length(df.s11.mean$citric.acid)
length(df.s12.mean$citric.acid)
length(df.s11.median$citric.acid)
length(df.s12.median$citric.acid)
```
From the table above, it appears there are outliers in the data as the graph suggests. The max value of 1 is more than 3 STDs from the mean

```{r echo=FALSE}
# descriptive stats for total.acidity
desc_stats.fun(df$total.acidity)

#subsetting data to count how many entries are above and below the mean and median
df.s13.mean <- subset(df, total.acidity > mean(total.acidity))
df.s14.mean <- subset(df, total.acidity < mean(total.acidity))
df.s13.median <- subset(df, total.acidity > median(total.acidity))
df.s14.median <- subset(df, total.acidity < median(total.acidity))
length(df.s13.mean$total.acidity)
length(df.s14.mean$total.acidity)
length(df.s13.median$total.acidity)
length(df.s14.median$total.acidity)
```

From the table above, it appears there are outliers in the data as the graph suggests. The max value of 17.045 is more than 3 STDs from the mean



```{r echo=FALSE}
# sulphates exploration
histogram.fun(df$sulphates,df) +
  xlab("Sulphates") +
  coord_cartesian(xlim = c(0.3,quantile(df$sulphates,.99)))
```
Sulphates has a normal distribution between about 0 and 1 g/ml and appear to have outliers above 1.5 g. Lets looks at some \
descriptive statistics for sulphates. Top 1% have been omitted from graph \

```{r echo=FALSE}
# descriptive stats for sulphates
desc_stats.fun(df$sulphates)

#subsetting data to count how many entries are above and below the mean and median
df.s5.mean <- subset(df, sulphates > mean(sulphates))
df.s6.mean <- subset(df, sulphates < mean(sulphates))
df.s5.median <- subset(df, sulphates > median(sulphates))
df.s6.median <- subset(df, sulphates < median(sulphates))


length(df.s5.mean$sulphates)
length(df.s6.mean$sulphates)
length(df.s5.median$sulphates)
length(df.s6.median$sulphates)
```

From the table above, it appears there are outliers in the data as the graph suggests. The max value of 2 is more than 3 STDs from the mean



> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis



### Structure of data set:
The data set  contains 1599 observations with 13 different variables. \
'data.frame':	1599 obs. of  14 variables:
 

### Main Feature of interest:
The main feature I am interested in from this data set is how wine Quality is affected. 

### Other features in the dataset that will help support \
investigation into feature of interest?
I will explore the correlation of wine quality with the other variables to help determine the other features I will use to support my \
investigation

### New Variables created:
The following new variables were created to help aid my data exploration in the bivariate and multivariate sections. 

* df$qaulity.bucket
* df$total.acidity

### Features Explored 

* Wine Quality
The wine Quality distribution looks like a normal distrubtion with a mean of `r mean(df$quality)`. Majority \
of the Quality ratings where above the mean. There were `r length(df.s2$quality)` above the mean and \
`r length(df.s1$quality)` below the mean. The actual distribution of quality ratings can be seen below

```{r echo=FALSE}
table(df$quality)
```
* Acidity \
There are two different types of acidity in wine, Fixed Acidity and Volatile Acidity. Volatile Acidity contributes to the overall taste \
of wine while non volatile wine can create a unpleasent vinegar taste with high quantities. The distributions for for fixed acidity and \
total acidity seem to be normal distrubtions while citric and volatile acidity look more bi-modal. From the graphs and tables for Acidity\
there are outliers present in every acidity category. To deal with the outliers, the best metric to explore is the median since it is \
more robust to outliers. 



* Sulphates \
From the summary tables for sulphates, we see that sulphates have a a min of .33 g/ml and a max of 2 g/ml. The mean for sulphates is .66. \
It appears that the max value for sulphates is an outlier due to it being over 3 STDs from the mean. There are are 964 wine rankings with \
sulphate levels below the mean and 635 rankings above the mean. In comparison, there are are 770 rankings above the median and \
768 rankings below the median. It appears the metric to best explore is the median across wine quality to help deal with outliers \ 
that are present in the data. 




# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

Before I begin the Bivariate exploration, I am going to load in the data set again and create factors out of the quality rankings \
to help with the supporting graphs. I will also add the total acidity variable in to the data frame to perform correlations against
wine quality rankings\

```{r echo=FALSE, Bivariate_Plots}
# Wine Quality correlation
df <- read.csv('wineQualityReds.csv')
df$total.acidity <- df$fixed.acidity + df$volatile.acidity + df$citric.acid
df$quality.bucket <- cut(df$quality, breaks = c('2','4','6','8'), labels = c("low","medium","high"))
c1 <- cor(x=df[,2:14], y= df$quality)
c2 <- cor(x=df[,2:14])
quality.factors <- factor(df$quality, levels=c(0,1,2,3,4,5,6,7,8,9,10))
df$quality.factors <- quality.factors
ggcorrplot(c1,  method = 'circle', lab = TRUE, lab_size = 3)
```

From the correlation matrix for Quality, I will explore the top 3 positive correlations to see the effect on quality. Those are: \
1. Alcohol \
2. Sulphates \
3. citric acid \ 

I am also interested in exploring the largest negative correlation to see what the effect on Quality is. The variable I will look at is \
1. Volatile Acidity


```{r echo=FALSE, BoxPlot_Function}
boxplot.fun <- function(df, xvar,yvar){
  plot <- ggplot(data = df, aes_string(x=xvar,y=yvar)) +
    geom_boxplot()
  return (plot)
}

```

```{r echo=FALSE, Line_Function}
line.fun <- function(df, xvar,yvar){
  plot <- ggplot(data = df, aes_string(x=xvar,y=yvar)) +
    geom_line(aes(group = 1),colour = 'blue') 
  return (plot)
}

```

```{r echo=FALSE, Scatter_Plot_Function}
scatterplot.fun <- function(df, xvar,yvar){
  plot <- ggplot(aes_string(x = xvar, y = yvar), data = df) +
    geom_jitter(color = 'blue', alpha = 1/10) +
    geom_line(stat = 'summary', fun.y = mean, color = 'green') +
    geom_line(stat = 'summary', fun.y = median, color = 'purple') +
    geom_smooth(method = 'lm', aes( group = 1), color = 'black')
  return (plot)
}

```

### Alcohol Exploration \

To start off, lets explore how the alcohol percentage is distributed accross wine quality to try and get an understanding of how \
alcohol relates to high quality wines

```{r echo=FALSE}
# Alcohol mean distributions across the different wine quality rankings. 
mean_df.alcohol <- df %>% group_by(quality.factors) %>%
  summarise(Alcohol_Mean = mean(alcohol),Alcohol_Median = median(alcohol), n = n())
mean_df.alcohol
```

As stated earlier in the univariate section, the mean wine quality was about 5.6% and there were more entries above the mean then below.\ 
In the \ summary table above, it appears that as wine quality rankings above 5% increase, so does the alcohol mean

Now lets see how this looks like visually on a box plot





#```{r echo = False}
# Alcohol vs Quality Box plot
#boxplot.fun(df,'quality.factors','alcohol') +
#  geom_hline(yintercept = mean(df$alcohol), linetype='longdash', alpha=.5, color='red') +
#  geom_hline(yintercept = median(df$alcohol), linetype='longdash', alpha=.5, color='orange')
  

#```


As seen above, the boxplot confirms that as the quality increases beyond the quality mean, the mean of alcohol seems to increase as well \
The red dotted line represents the mean for alcohol and the yellow dotted line represents the median. \
When we graph the alcohol means, we can see a similar pattern with the median as well. See graph below

```{r echo=FALSE}
# Alcohol vs Quality Median Line Graphs
p6 <- line.fun(mean_df.alcohol,'quality.factors','Alcohol_Mean')
p7 <- line.fun(mean_df.alcohol,'quality.factors','Alcohol_Median')
grid.arrange(p6,p7, ncol=2)
```

As seen above, as the Wine Quality increases, the alcohol percentage mean and medians increase as well at around a ranking of 5. 



Lets see how this looks on a scatter plot.


```{r echo=False}
#Alcohol Vs Quality Scatter Plot
scatterplot.fun(df,'quality','alcohol') +
    geom_hline(yintercept = mean(df$alcohol), linetype='longdash', alpha=.5, color='red') +
    geom_hline(yintercept = median(df$alcohol), linetype='longdash', alpha=.5, color='orange')
```



From the scatter plot, we can see that majority of the high quality wines have alcohol percentages that are above the mean and medians. \
In contrast, we see majority of the lower quality wines have alcohol percentages below the mean and medians. 

### Sulphates Exploration \

The next wine property I will explore is Sulphates and see how it relates to Wine Quality Rankings. Lets remember that Sulphates \
are not as strongly correlated to quality as Alcohol. 

To start off, lets explore how sulfates is distributed accross wine quality to try and get an understanding of how sulphates relate \
to high quality wines

```{r echo=FALSE}
# Sulphates mean and median distributions across the different wine quality rankings. 
mean_df.sulphates <- df %>% group_by(quality) %>%
  summarise(sulphates_mean = mean(sulphates),sulphates_median = median(sulphates), n = n())
mean_df.sulphates

```
As stated in the univariate analysis, the metric I will explore is the median to deal with the outliers. Sulfates \
 have a median of .62 g/ml. It appears that as wine quality rankings increase, sulphate levels also increase. \
 Lets looks at this visually.

```{r echo=FALSE}
boxplot.fun(df,'quality.factors','sulphates') +
  geom_hline(yintercept = mean(df$sulphates), linetype='longdash', alpha=.5, color='red') +
  geom_hline(yintercept = median(df$sulphates), linetype='longdash', alpha=.5, color='orange')
```


As seen above in the box plot, it appears higher quality wines have higher levels of sulphates that are \
above the mean and median. Lets see how the median levels look like.

```{r echo = False}
# Sulphates vs Quality median line graph
line.fun(mean_df.sulphates,'quality','sulphates_median') +
  scale_y_continuous(limits = c(.5, .75), breaks = seq(.5,.8,.05)) +
  geom_hline(yintercept = median(df$sulphates), linetype='longdash', alpha=1, color='orange') +
  geom_hline(yintercept = mean(df$sulphates), linetype='longdash', alpha=1, color='red')
```



The graph above confirm our obversations seem in the tables that as the wine quality increases, the sulphates level increase \
It appears that the high quality wines contain a sulphate level that is above both the median and mean levels of sulphate. \
The yellow dotted line is the median and the red dotted line is the mean. 

Lets see how this looks on a scatterplot 

```{r echo = False}
# Sulphates vs Quality Scatter plot
scatterplot.fun(df,'quality','sulphates') +
    geom_hline(yintercept = mean(df$sulphates), linetype='longdash', alpha=.5, color='red') +
    geom_hline(yintercept = median(df$sulphates), linetype='longdash', alpha=.5, color='orange') +
    coord_cartesian( ylim = c(.4,quantile(df$sulphates,.99)))
```




From the scatter plot, the top 1% have been omitted to deal with outliers and we can see that there are \
fewer high quality wines with sulphate levels below the mean and median than above it. \
Additionally we can see that there are more lower quality wines that have sulphate levels below the mean and median levels than above. 

### Citric Acid Exploration \

Moving on from the Sulphates exploration, I am now going to explore how citric acid relates to Wine Quality Rankings \
First, lets remember that citric acid is not as strongly correlated to quality as Alcohol percentage.  

To start off, lets explore how citric acid is distributed accross wine quality to try and get an understanding of citric acid relates \
to high quality wines


```{r echo=FALSE}
# citric acid mean distributions across the different wine quality rankings. 
mean_df.citric_acid <- df %>% group_by(quality) %>%
  summarise(citric_acid_mean = mean(citric.acid),citric_acid_median = median(citric.acid), n = n())
mean_df.citric_acid
```

In the case for citric acid, it looks like it follows a similar behaviour as sulphates. \
As stated in the univariate analysis, the metric I will explore is the median to deal with the outliers. There seems to be nore \
outliers in citric acid with the big difference in the means and medians found in the lower quality wines. The median for citric acid \
is .26 g/ml. An interesting obervation is that there are more citric acid levels below both the mean and median in contrast to sulphates.\

Lets looks at this visually. \

```{r echo=FALSE}
#Box plot of Citric Acid means by Wine Quality Rankings
boxplot.fun(df,'quality.factors','citric.acid') +
  geom_hline(yintercept = mean(df$citric.acid), linetype='longdash', alpha=.5, color='red') +
  geom_hline(yintercept = median(df$citric.acid), linetype='longdash', alpha=.5, color='orange')

```



The box plot above confirms the obersvations made from the median table. It appears the higher quality wines with rankings of 7 and 8 \
have citric acid levels above the mean and the median. \

Lets take a visual look at the medians for citric acid across the different wine quality rankings. \

```{r echo=FALSE}
# Plot for citric acid medians acorss the quality rankings. 
line.fun(mean_df.citric_acid,'quality','citric_acid_median') +
    geom_hline(yintercept = median(df$citric.acid), linetype='longdash', alpha=1, color='orange') +
    geom_hline(yintercept = mean(df$citric.acid), linetype='longdash', alpha=1, color='red') +
    scale_x_continuous(limits = c(3, 8), breaks = seq(3,8,1)) 
```



The graph above also confirms the observation that as wine quality increases, citric acid levels incease as well. Additionally, \
similar to sulphates, high quality wines have citric acid levels that are above both the mean and medians


Finally, lets take a look at a scatter plot 

```{r echo=FALSE}
scatterplot.fun(df,'quality','citric.acid') +
    geom_hline(yintercept = mean(df$citric.acid), linetype='longdash', alpha=.5, color='red') +
    geom_hline(yintercept = median(df$citric.acid), linetype='longdash', alpha=.5, color='orange') +
    coord_cartesian( ylim = c(0,quantile(df$citric.acid,.99)))
```



From the scatter plot, the top 1% have been omitted to deal with outliers. We can see that majority of the high quality wines \
are have citric levels above the mean and median. However we can also see that there a more lower quality wines \
that are above the mean and median as well. This shows that citric acid is not very strongly correlated to quality. 


### Volatile Acidity \
Now that we have a better idea of the properties that make a good wine, lets explore a property that makes a bad wine. \

First, lets remember that volatile acidity is negatively correlated to quality.\

To start off, lets explore how volatile acidity is distributed accross wine quality to try and get an understanding of 
how volatile acidity relates to wine quality. \

```{r echo=FALSE}
summary_df.volatile_acidity <- df %>% group_by(quality) %>%
  summarise(volatile_acidity_mean = mean(volatile.acidity),volatile_acidity_median = median(volatile.acidity), n = n())
summary_df.volatile_acidity
```
As stated in the bivariate anaylsis, i will be exploring the median metric to deal with outliers in the data \
Looking at the mean/median table above, it is interesting to note that as the wine quality increases, volatile acidity levels decrease. \
This has to do with the negative correlation. \

Lets see how this looks visually. \

```{r echo=FALSE}
# Volatile Acidity vs Quality Box plot
boxplot.fun(df,'quality.factors','volatile.acidity') +
  geom_hline(yintercept = mean(df$volatile.acidity), linetype='longdash', alpha=.5, color='red') +
  geom_hline(yintercept = median(df$volatile.acidity), linetype='longdash', alpha=.5, color='orange')
```


The box plot above seems to support our observation that the lowest quality wines contains the most volatile acidity levels. It is \
important to note that those levels are above the mean and median. The high quality wines have volatile acidity levels well below the \
mean and medians. \

Lets take a look at the means and medians from each wine ranking. \

```{r echo=FALSE}
# Volatiele acidity median across the wine rankings
line.fun(summary_df.volatile_acidity,'quality','volatile_acidity_median') +
  geom_hline(yintercept = median(df$volatile.acidity), linetype='longdash', alpha=1, color='orange') +
  geom_hline(yintercept = mean(df$volatile.acidity), linetype='longdash', alpha=1, color='red') +
  scale_x_continuous(limits = c(3, 8), breaks = seq(3,8,1)) 
```


Again, we can see that the lower quality wines  have medians that are significantly higer than the higher quality wines. \

Finally lets take a look at a scatter plot. \

```{r echo=FALSE}
scatterplot.fun(df,'quality','volatile.acidity') +
    geom_hline(yintercept = mean(df$volatile.acidity), linetype='longdash', alpha=.5, color='red') +
    geom_hline(yintercept = median(df$volatile.acidity), linetype='longdash', alpha=.5, color='orange') +
    coord_cartesian( ylim = c(0.1,quantile(df$volatile.acidity,.99)))
```


From the scatter plot, we see that there are very few low quality wines with volatile acidity levels that are below the mean and median levels\
Conversly, we that there are few high quality wines that have volatile acidity levels above the mean and median levels. \


```{r echo=FALSE}
# Correlation accross all 14 variables. 
ggcorrplot(c2, type = "lower", method = 'square', lab = TRUE, lab_size = 2.5,hc.order = TRUE)
```



# Bivariate Analysis \

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.


### My eploration consited of how different properties affect overal wine quality. \

### Alcohol Percentage vs Quality \

The alcohol percentage had the strongest relationship for wine quality than any other property with a co-efficient value of .48. \
I discovered that as the alcohol percentage increased, the quality seemed to increase. The higher quality wines rated 7 and 8 \
had much higher alcohol means and medians than the lower quality wines. As seen from the scatter plot, there are more high quality wines \
that have alcohol percentages above the alcohol means and medians than below it. 

### Sulphate levels vs Quality \

The correlation between sulphate and quality was the 2nd strongest relationship that carried a co-efficient of .25. \
Sulphate seemed to experience a similar pattern as the alcohol percentage did. As wine quality increased, sulphate levels increased. \
The higher quality wines had high sulphate means and medians that the lower quality wines. As seen from the scatter plot, there are \
very few high quality wines that have sulphate levels below the mean and medians. However, due to the weaker relationship, there appears \
to be a good amount of lower quality wines that are above the mean and median sulphate levels. 

### Citric Acid levels vs Quality \

The correlation between citric acid and quality was the 3rd strongest relationship that carried a co-efficient of .23. \
Citric Acid has a similar relationship to sulphates and Alcohol. As the wine quality increased, the citric acid means and medians \
increased as well. Majority of the high quality wines had citric acid levels above the mean and median levels. Similar to sulphates and \
due to the weaker relationship on quality, there were a good amount of lower quality wines above the mean and median levels. 

### Volatile Acidity vs Quality \

My last exploration was on Volatile acidity vs Quality. This relationship had the strongest negative relationship with a co-efficient \
of -.39. It was interesting to see that volatile acidity had an opposite affect on quality. As volatile acidity decreased, wine quality \
increased as suggested by the correlation relationship. As the wine quality increased, the volatile acidity mean and median levels \
decreased. As seen in the scatter plot there are very few high quality wines that have volatile acidity levels above the mean and median \
levels. 



# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.


In the Bivariate section, we saw how quality was affected by alcohol percentage,citric acid, sulphates, and volatile acidity. \
I am now going to explore the relationship between these properties across the differene quality buckets. \
All graphs will contain data in the 1-99% quantile range to omit any outliers in the data and will be split into 4 quandrants by the means\
and medians from each property explored. 


```{r echo = False, message=FALSE, warning=FALSE,Multiplot_Function}
#Multiplot Function
multiplot.fun <- function(df, xvar,yvar){
  plot <- ggplot(data = df, aes_string(x = xvar, y=yvar)) +
    geom_jitter(alpha=.4, aes(size=quality.bucket, color=quality.bucket)) +
    theme_dark() +
    scale_color_brewer(palette =  'Paired') +
    geom_smooth(method = "lm", se = FALSE,size=1, color = 'red')
    
  return(plot)
}
```

The first relationship I will explore is alcohol vs citric acid. \

```{r echo = False, message=FALSE, warning=FALSE,Multiplot_1}
multiplot.fun(df,'alcohol','citric.acid') +
  geom_vline(xintercept = median(df$alcohol), linetype = 'longdash', color = 'black') +
  geom_hline(yintercept = median(df$citric.acid),linetype = 'longdash', color = 'black') +
  geom_vline(xintercept = mean(df$alcohol), linetype = 'longdash', color = 'red') +
  geom_hline(yintercept = mean(df$citric.acid),linetype = 'longdash', color = 'red') +
  coord_cartesian(xlim=c(quantile(df$alcohol,.01),quantile(df$alcohol,.99)),ylim=c(quantile(df$citric.acid,.01),quantile(df$citric.acid,.99))) 
  
```






In the graph above displays data in the 1-99% quantile to omit any outiers. The wine quality rankings were grouped into 3 different groups. \
* 3-4 = Low \
* 5-6 = Medium \
* 7-8 = High \


We can see that the majority of the high quality wines are above the means and medians of the alcohol percentage and citric acid levels. \
It appears that the highest quality wines of 7 and 8, have a combination of citric acid levels of at least \
`r median(df$citric.acid)`g/ml and alcohol percentage of `r median(df$alcohol)`% \

Next I would like explore the relation between sulphate and citric acid across the different wine qualities. As seen in the correlation \
table in the bivariate section, sulphate and citric acid have a correlation co-efficient of .31 and we already saw that these two \
properties have an affect on quality. 

```{r echo = False, message=FALSE, warning=FALSE,Multiplot_2}
multiplot.fun(df,'sulphates','citric.acid') +
  geom_vline(xintercept = median(df$sulphates), linetype = 'longdash', color = 'black') +
  geom_hline(yintercept = median(df$citric.acid),linetype = 'longdash', color = 'black') +
  geom_vline(xintercept = mean(df$sulphates), linetype = 'longdash', color = 'red') +
  geom_hline(yintercept = mean(df$citric.acid),linetype = 'longdash', color = 'red') +
  coord_cartesian(xlim=c(quantile(df$sulphates,.01),quantile(df$sulphates,.99)),ylim=c(quantile(df$citric.acid,.01),quantile(df$citric.acid,.99)))
```

The graph above tells a very similar story when we graphed alcohol vs citric acid across the different quality buckets. Like the 1st graph, \
the graph displays data from the 1-99% quantiles to remove outliers. It appears that majority of wines ranked 7 or 8 have \
a combination of sulphate leve;s that are at least `r median(df$sulphates)`g/ml and \
citric acid levels of at least `r median(df$citric.acid)` g/ml


The next relation i will look at is Alcohol vs Sulphates and explore the relation to quality.\ 


```{r echo = False, message=FALSE, warning=FALSE,Multiplot_3}
multiplot.fun(df,'alcohol','sulphates') +
  geom_vline(xintercept = median(df$alcohol), linetype = 'longdash', color = 'black') +
  geom_hline(yintercept = median(df$sulphates),linetype = 'longdash', color = 'black') +
  geom_vline(xintercept = mean(df$alcohol), linetype = 'longdash', color = 'red') +
  geom_hline(yintercept = mean(df$sulphates),linetype = 'longdash', color = 'red') +
  coord_cartesian(xlim=c(quantile(df$alcohol,.01),quantile(df$alcohol,.99)),ylim=c(quantile(df$sulphates,.01),quantile(df$sulphates,.99))) 
  
```

In this third graph, we can begin to see a consistent pattern in the top 3 positively correlated properties on quality. The top wine rated \
7 and 8 have a combination of alochol percentage of at least `r median(df$alcohol)`% and `r median(df$sulphates)` g/ml \

Now I am going to explore the combination of Volatile Acidy with Alcohol,Citric Acid,and Sulphates to see the effect on quality. \
I'll first start with Alcohol vs Volatile Acidity. 

```{r echo = False, message=FALSE, warning=FALSE,Multiplot_4}
multiplot.fun(df,'alcohol','volatile.acidity') +
  geom_vline(xintercept = median(df$alcohol), linetype = 'longdash', color = 'black') +
  geom_hline(yintercept = median(df$volatile.acidity),linetype = 'longdash', color = 'black') +
  geom_vline(xintercept = mean(df$alcohol), linetype = 'longdash', color = 'red') +
  geom_hline(yintercept = mean(df$volatile.acidity),linetype = 'longdash', color = 'red') +
  coord_cartesian(xlim=c(quantile(df$alcohol,.01),quantile(df$alcohol,.99)),
                  ylim=c(quantile(df$volatile.acidity,.01),quantile(df$volatile.acidity,.99))) 

```
We can clearly see that majority of the wines rated 7 and 8 have an alcohol percentage of at least `r median(df$alcohol)`% and \
 have volatile acidity levels less than `r median(df$volatile.acidity)` g/ml \
 
 Next up is Citric acid vs Volatile Acid \
 
 
```{r echo = False, message=FALSE, warning=FALSE,Multiplot_5}
multiplot.fun(df,'citric.acid','volatile.acidity') +
  geom_vline(xintercept = median(df$citric.acid), linetype = 'longdash', color = 'black') +
  geom_hline(yintercept = median(df$volatile.acidity),linetype = 'longdash', color = 'black') +
  geom_vline(xintercept = mean(df$citric.acid), linetype = 'longdash', color = 'red') +
  geom_hline(yintercept = mean(df$volatile.acidity),linetype = 'longdash', color = 'red') +
  coord_cartesian(xlim=c(quantile(df$citric.acid,.01),quantile(df$citric.acid,.99)),
                  ylim=c(quantile(df$volatile.acidity,.01),quantile(df$volatile.acidity,.99))) 

```
In the graph above, we see majority of the wines rated between 7 and 8 are all in the lower right quadrant of the graph which is very similar\
to the alcohol vs volatile acidity graph. We see that majority of these wines have no more than `r median(df$volatile.acidity)` g/ml \
of volatile acidity and at least `r median(df$citric.acid)` g/ml of citric acid\

The final reltionship I'll explore is volatile acidity vs sulphates. 

```{r echo = False, message=FALSE, warning=FALSE,Multiplot_6}
multiplot.fun(df,'sulphates','volatile.acidity') +
  geom_vline(xintercept = median(df$sulphates), linetype = 'longdash', color = 'black') +
  geom_hline(yintercept = median(df$volatile.acidity),linetype = 'longdash', color = 'black') +
  geom_vline(xintercept = mean(df$sulphates), linetype = 'longdash', color = 'red') +
  geom_hline(yintercept = mean(df$volatile.acidity),linetype = 'longdash', color = 'red') +
  coord_cartesian(xlim=c(quantile(df$sulphates,.01),quantile(df$sulphates,.99)),
                  ylim=c(quantile(df$volatile.acidity,.01),quantile(df$volatile.acidity,.99))) 

```
From the Final graph above, we can see the same pattern from the previous 2 graphs. Majority of the high quality wines are in the bottom \
right quandrant of the graph. High quality wines at no more than `r median(df$volatile.acidity)` g/ml \
of volatile acidity and at least `r median(df$sulphates)` g/ml of sulphates\

# Multivariate Analysis



After exploring the combination affects of Alcohol,citric acid,sulphates,and volatile acidity on quality, some interesting observations can \
be seen. \
1. As seen in the bivariate section, Alcohol,Citric Acid,and Sulphates are all positively correlated to Quality. This relationship shows up \
in the multivariate plots. When you graph the combination affect of these properties, one can notice that majority of high quality wines \
are in the upper right quadrant and above the means and medians for property. However, The combination affect of these properties \
are not as strong for the medium and low quality wines as they are spread across all the quandrants of the graphs \

2. As seen in the bivariate section, volatile acidity was negatively correlated to quality and to the other properties I mentioned earlier \
When you graph the combination affect of volatile acidity with the other 3 properties across the quality buckets, you can see this relationship\
in the high quality wines. Majority of the high quality wines are in the lower right quadrants of the graph and had volatile acidity levels \
below the median of volatile acidity. However, this also is not true for the medium and low quality wines. It appears the high quality wines \
are most affected by this correlation relationship. 

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
