---
output:
  html_document: default
  pdf_document: default
---
 Wine by Nick Olson
=====================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, Packages}

# Install packages and libraries
#install.packages("ggplot2")
#install.packages("ggcorrplot")

library(ggplot2)
library(ggcorrplot)
library(gridExtra)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_data}

#Find where the dataset is and write to variable

setwd("C:/Users/nolson8.COLLEGES/Desktop/Python/R")
wine <- read.csv("wineQualityReds.csv", sep = ',', header=T)

```

I will explore a wine dataset that focuses on the characteristics of red wine.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Str}

#View the structure of the dataset

str(wine)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Summary}

#Summary of the variable

summary(wine)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Dim}

#Find how many observations and variables are in the dataset

dim(wine)

```

The dataset has 13 variables and 1599 observations

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality1}

#Write a new variable that groups quality

wine$quality1 <- cut(wine$quality, c(2,4,6,8),
                     labels = c("Low", "Medium", "High"))

```

Observations are grouped into quality ratings 3-8 and then ranked in to 
categories Low, Medium and High

```{r echo=FALSE, message=FALSE, warning=FALSE, Table}

#View tables of variables

table(wine$quality)

table(wine$quality1)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Function1}

#Create a function

graphing.quality <- function(quality, binwidth = 0.01) {
  return(ggplot(aes_string(x = quality), data = wine) + 
           geom_histogram(stat = "count"))
  }

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots1}

#Create variables for graphs and arrange into a grid


q.plot1 <- graphing.quality('quality1',0.01)

q.plot2 <- graphing.quality('quality',0.01)

grid.arrange(q.plot1, q.plot2, ncol=2)

```

After plotting the Red Wine Dataset most wines have a quality of 5 or 6 while 
low quality and high quality are less common.  When I created a new variable 
that groups quality into 3 categories, 'Medium' quality had the most 
observations.

Next I wanted to look at the characteristics (input variables) of the red wine.

```{r echo=FALSE, message=FALSE, warning=FALSE, Function2}

#Create function for histograms

hist.graph <- function(wine, xvar_string) {
  wine1 <- ggplot(data=wine, aes_string(x=xvar_string)) +
    geom_histogram(bins = 100)
  wine1
}

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, Plots2}

#Create new variables to graph function. Arrange into grid.


alcohol.plot1 <- hist.graph(wine, "alcohol")

sulphates.plot1 <- hist.graph(wine, "sulphates")

density.plot1 <- hist.graph(wine, "density")

total.sulfur.dioxide.plot1 <- hist.graph(wine, "total.sulfur.dioxide")

free.sulfur.dioxide.plot1 <- hist.graph(wine, "free.sulfur.dioxide")

chlorides.plot1 <- hist.graph(wine, "chlorides")

residual.sugar.plot1 <- hist.graph(wine, "residual.sugar")

citric.acid.plot1 <- hist.graph(wine, "citric.acid")

volatile.acidity.plot1 <- hist.graph(wine, "volatile.acidity")

fixed.acidity.plot1 <- hist.graph(wine, "fixed.acidity")

pH.plot1 <- hist.graph(wine, "pH")

grid.arrange(alcohol.plot1, sulphates.plot1, density.plot1, 
             total.sulfur.dioxide.plot1, free.sulfur.dioxide.plot1, 
             chlorides.plot1, residual.sugar.plot1, citric.acid.plot1, 
             volatile.acidity.plot1, fixed.acidity.plot1, pH.plot1, ncol = 3)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, summarySugar}

summary(wine$residual.sugar)

```

While some of the graphs had normal distribution, some of them were skewed to 
the left. Others were bimodal. I also noticed that Citric Acid had two peaks at 
0 and .49.  Residual Sugar and Chlorides had some of the biggest outliers.
Residual Sugar had a mean of 2.539 however, it had a max of 15.5.  

```{r echo=FALSE, message=FALSE, warning=FALSE, logscale}

sulphates.log <- qplot(x = log10(sulphates), data = wine, bins = 100)

total.sulfur.dioxide.log <- qplot(x = log10(total.sulfur.dioxide), 
                                  data = wine, bins = 100)

free.sulfur.dioxide.log <- qplot(x = log10(free.sulfur.dioxide), 
                                 data = wine, bins = 100)

chlorides.log <- qplot(x = log10(chlorides), data = wine, bins = 100)

residual.sugar.log <- qplot(x = residual.sugar, data = wine, bins = 100)

grid.arrange(sulphates.log, total.sulfur.dioxide.log, free.sulfur.dioxide.log, 
             chlorides.log, residual.sugar.log, ncol = 2)

```

Some of the features had outliers and were not close to a normal distribution. 
I took the log10 of those features to get a more normal distribution.

# Univariate Analysis

### What is the structure of your dataset?

There are 1599 wine observations in the dataset.  The dataset included 12 
features (fixed acidity, volatile acidity, citric acid, residual sugar, 
chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, 
alcohol, quality).  
All of the features are numbered (except quality which is an integer). 
None of them are order factored variables.

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest to me is the quality of the wine. I would like to 
see which other features effect the quality of the wine.  

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Based on my current knowledge of wines. I am not sure which variables will have 
any correlation with the quality.  However, based on the plots I would think 
that density and/or the citric acid would affect the quality. 

### Did you create any new variables from existing variables in the dataset?

I created a new variable for wine quality.  I grouped the observations in "Low",
"Medium" and "High" quality.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

After increasing the bin size for citirc acid I noticed that quite a few wines 
have either 0 g/dm^3 or .49 g/dm^3 of citric acid.

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation}

#Create a variable that holds a correlation matrix

winedata <- wine[, c(2,3,4,5,6,7,8,9,10,11,12,13)]

corr <- cor(winedata)
corr

ggcorrplot(corr, type = "lower", method = 'square', lab = TRUE, lab_size = 2.5,
           hc.order = TRUE)

```

This correlation matrix indicates that there is a positive correlation between 
quality and alcohol, citric acid, sulphates and fixed acidity.  It also shows a 
negative correlation between volatile acidity, total sulfur dioxide, chlorides 
and density.

-Quality of wine increasing
```{r echo=FALSE, message=FALSE, warning=FALSE, Function3}

box.plot.graphs <- function(wine, xvar_string, yvar_string){
  gg.graph <- ggplot(aes_string(x = xvar_string, y = yvar_string), 
                     data = wine) +
    geom_jitter( alpha = .3)  +
    geom_boxplot( alpha = .5,color = 'blue')+
    stat_summary(fun.y = "mean", geom = "point", color = "red",
                 shape = 8, size = 4) 
  gg.graph
  }

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots3}

#Create variables that hold graphs of positive correlation

positive.graph1 <- box.plot.graphs(wine, 'quality1', 'citric.acid')

positive.graph2 <- box.plot.graphs(wine, 'quality1', 'sulphates')

positive.graph3 <- box.plot.graphs(wine, 'quality1', 'fixed.acidity')

positive.graph4 <- box.plot.graphs(wine, 'quality1', 'alcohol')

grid.arrange(positive.graph1, positive.graph2, positive.graph3, positive.graph4,
             ncol = 2)

```

I chose to analyze Alcohol, Citric Acid, Sulphates and Fixed Acidity because 
they were the top positively correlated features with quality. 
Alcohol had the biggest correlation with quality (0.48). It looks as though 
alcohol content is greater in 'High' quality wines.  All of these features seem
to increase along with the quality.

-Quality of Wine decreasing

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots4}

#Create variables that hold graphs of negative correlation

negative.graph1 <- box.plot.graphs(wine, 'quality1', 'volatile.acidity')

negative.graph2 <- box.plot.graphs(wine, 'quality1', 'total.sulfur.dioxide')

negative.graph3 <- box.plot.graphs(wine, 'quality1', 'density')

negative.graph4 <- box.plot.graphs(wine, 'quality1', 'chlorides')

grid.arrange(negative.graph1, negative.graph2, negative.graph3, negative.graph4,
             ncol = 2)

```

Volatile Acidity, Total Sulfur Dioxide, Density and Chlorides were the features 
with the largest negative correlations with the wines quality. Volatile Acidity 
had the biggest negative correlation with quality (-0.39). Volatile Acidity was 
much higher in 'Low' quality wines. As you can tell from the graphs the negative
correlation for the other 3 features was small.

-Other Correlations

Positive Correlations

```{r echo=FALSE, message=FALSE, warning=FALSE, Function4}

corr.graphs <- function(wine, xvar_string, yvar_string){
  gg.graph <- ggplot(aes_string(x = xvar_string, y = yvar_string, 
                                color ='quality1'), 
                      data = wine) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(method = "lm", se = FALSE,size=1)  +
    scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) + 
    theme_dark() 
  gg.graph
  }

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots5}

#Create variable that hold graphs of positive correlation

corr.graph1 <- corr.graphs(wine, 'free.sulfur.dioxide', 'total.sulfur.dioxide')

corr.log.graph1 <- ggplot(aes(x = log10(total.sulfur.dioxide), 
                 y = log10(free.sulfur.dioxide), colour = quality1), 
             data = wine) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(method = "lm", se = FALSE,size=1)  +
    scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) + 
    theme_dark() +
  scale_x_continuous(lim=c(quantile(log10(wine$total.sulfur.dioxide),
                                    0.01),
                           quantile(log10(wine$total.sulfur.dioxide),
                                    0.99))) +
  scale_y_continuous(lim=c(quantile(log10(wine$free.sulfur.dioxide),
                                    0.01),
                           quantile(log10(wine$free.sulfur.dioxide),
                                    0.99)))

grid.arrange(corr.graph1, corr.log.graph1, ncol = 1)


```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots6}

#Create variable that hold graphs of positive correlation

corr.graph2 <- corr.graphs(wine, 'density', 'fixed.acidity')

corr.graph2

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots7}

#Create variable that hold graphs of positive correlation

corr.graph3 <- corr.graphs(wine, 'fixed.acidity', 'citric.acid')

corr.graph3

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots8}

#Create variable that hold graphs of positive correlation

corr.graph4 <- corr.graphs(wine, 'sulphates', 'alcohol')

corr.log.graph4 <- ggplot(aes(x = log10(sulphates), 
                 y = alcohol, colour = quality1), 
             data = wine) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(method = "lm", se = FALSE,size=1)  +
    scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) + 
    theme_dark() +
  scale_x_continuous(lim=c(quantile(log10(wine$sulphates),
                                    0.01),
                           quantile(log10(wine$sulphates),
                                    0.99)))

grid.arrange(corr.graph4, corr.log.graph4, ncol = 1)

```

I chose to graph 3 features the greatest positive correlations. I also wanted to
graph the correlation between Alcohol and Sulfates. These two features were not 
strongly correlated with each other. However, they were the top 2 features that 
were positivly correlated with the quality of the wine. As you can see, most 
observations have less than 1 g/dm3 of sulphates. I also wanted to log transform
the graphs that were not normal distribution. I log transfored the features 
Free Sulfur Dioxide and Total Sulfur Dioxide.

Negative Correlations

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots9}

#Create variable that hold graphs of negative correlation

corr.graph5 <- corr.graphs(wine, 'fixed.acidity', 'pH')

corr.graph5

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots10}

#Create variable that hold graphs of negative correlation

corr.graph6 <- corr.graphs(wine, 'citric.acid', 'volatile.acidity')

corr.graph6

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots11}

#Create variable that hold graphs of negative correlation

corr.graph7 <- corr.graphs(wine, 'density', 'alcohol')

corr.graph7

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Plots12}

#Create variable that hold graphs of negative correlation

corr.graph8 <- corr.graphs(wine, 'volatile.acidity', 'total.sulfur.dioxide')

corr.log.graph8 <- ggplot(aes(x = volatile.acidity, 
                 y = log10(total.sulfur.dioxide), colour = quality1), 
             data = wine) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(method = "lm", se = FALSE,size=1)  +
    scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) + 
    theme_dark() +
  scale_y_continuous(lim=c(quantile(log10(wine$free.sulfur.dioxide),
                                    0.01),
                           quantile(log10(wine$free.sulfur.dioxide),
                                    0.99)))

grid.arrange(corr.graph8, corr.log.graph8, ncol = 1)

```

These are graphs of the 3 features with the greatest negative correlations. I 
also wanted to graph the correlation between Total Sulfur Dioxide and Volatile 
Acidity. These two features were not strongly correlated with each other. In 
fact they had a positive correlation with each other. However, they were the top
2 features that were negatively correlated with the quality of the wine. I log 
transformed Total Sulfur Dioxide to account for outliers.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features \
in the dataset?

I noticed a few positive and negative correlations within the dataset.  
My main feature of interest was the observed quality of the wine.  The quality 
of wine generally went up when there was more citric acid present.  The opposite 
effect happend when there was more volatile acidity.  The more volatile acidity, 
the lower quality the wine.  This makes sense as both were negativly correlated.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

I noticed that there were a few features in the dataset that were either 
negatively or positively correlated.  Most of them were what I would have 
expected. I also looked at the relationship between alcohol and sulphates. 
Although they do not have a high positive correlation with eachother, they are 
both positively correlated with qulaity.

### What was the strongest relationship you found?

The strongest relationship that I found was the negative correlation between 
fixed acidity and pH.  

# Multivariate Plots Section

I chose to look at some graphs of features that were positively and negatively 
correlated

```{r echo=FALSE, message=FALSE, warning=FALSE, Function5}

multi.graphs <- function(wine, xvar_string, yvar_string){
  gg.graph <- ggplot(aes_string(x = xvar_string, y = yvar_string, 
                                color ='quality1'), 
                      data = wine) +
    geom_point() 
  gg.graph
  }

```

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph1}

multi.graph1 <- multi.graphs(wine, 'fixed.acidity', 'citric.acid')

multi.graph1

```

Citric Acid and Fixed Acidity were positively correlated (0.67).  It looks as if
the higher quality observations are positioned towards the top while a majority 
of lower quality observations are towards the bottom.  There are low quality 
observations that are scattered thoughout as well.

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph2}

multi.graph2 <- multi.graphs(wine, 'fixed.acidity', 'density')

multi.graph2

```

Density and Fixed Acidity are also positively correlated (0.67).  It looks like 
a majority of the high quality observations are located to the lower right side.
The medium and lower quality observations tend to fall towards the upper left 
side.

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph3}

multi.graph3 <- multi.graphs(wine, 'free.sulfur.dioxide', 'total.sulfur.dioxide')

multi.graph3

```

Total Sulfur Dioxide and Free Sulfur Dioxide are positively correlated as well 
(0.67). High quality observations look to fall on the right side while the lower
and medium quality seem to lean to the left side.

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph7}

multi.graph7 <- multi.graphs(wine, 'sulphates', 'alcohol')

multi.graph7

```

Alcohol and Sulphates were the most positively correlated features with quality.
They have a weak positive correlation with each other though. The graph 
indicates that most high quality observations have a higher alcohol content but 
not necessarily more sulphates.

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph4}

multi.graph4 <- multi.graphs(wine, 'fixed.acidity', 'pH')

multi.graph4

```

pH and Fixed Acidity are negatively correlated (-0.68). This is actually the 
biggest correlation between any features. It looks like the lower quality 
observations seem to be towards the top. You can see the one outlier near the 
bottom however.

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph5}

multi.graph5 <- multi.graphs(wine, 'citric.acid', 'volatile.acidity')
multi.graph5

```

Volatile Acidity and Citric Acid were also negatively correlated (-.55). Higher 
quality observations are towards the bottom right while lower quality are at the
top left.

```{r echo=FALSE, message=FALSE, warning=FALSE, multi.graph6}

multi.graph6 <- multi.graphs(wine, 'density', 'alcohol')

multi.graph6

```

Alcohol and Desity are negatively correlated (-0.55). High quality observations 
are towards the top left. Medium and lower quality are located towards the 
bottom.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of\
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

Because of the positive correlations between fixed acidity/citric acid, fixed 
acidity/density and free sulfur dioxide/total sulfur dioxide I decided to make 
multivariate plots showing the impact on quality.  Although the graphs are a 
little messy, you can see that qualities are grouped together. When looking at 
citric acid/fixed acidity's effects on quality you can see that high quality 
observations are grouped towards the top while medium is below it.  The majoriy 
of low quality appear at the bottom as well.


------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

#Graph quality

ggplot(wine, aes(quality1)) + geom_bar() + 
  labs(x = "Quality", y = "# of Observations") + 
  ggtitle("Red Wine Quality") + 
  theme(plot.title = element_text(hjust = 0.5))
                       

```

### Description One

There is a large number of observations that can be grouped into 'Medium' 
quality wine.  It is more rare to find wines that are 'Low' quality or 'High' 
quality.

### Plot Two

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

#Graph Quality and Alcohol

ggplot(aes(factor(quality1), alcohol), data = wine) +
  labs(x = "Quality", y = "Alcohol (% by Volume)")+ geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", geom = "point", color = "red", 
               shape = 8, size = 4) + ggtitle("Quality of Alcohol") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### Description Two

Alcohol had a positive correlation with quality.  Higher alcohol content seemed 
to have more observations that were grouped into 'High' quality wine. This graph 
is significant because it is the best indicator of the quality of wine.

### Plot Three

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

#Graph sulphates/alcohol and quality

ggplot(wine, aes(x = sulphates, y= alcohol, 
                                 color = quality1)) + geom_point() +
  ggtitle("Biggest Quality Factors Correlated") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(colour = "Quality",
       x = "Sulphates (g/dm3) ", y = "Alcohol (% by Volume)")

```

### Description Three

When you look at alcohol/sulphates relationship with quality, you can see that 
the points are grouped together. With the exception of outliers most of the low 
quality observations were on the left side of the graph. The higher quality 
observations seem to be higher and to the right.

------

# Reflection

The Red Wine dataset that I explored contained 1599 observations of quality.  
Looking into the dataset I found that interesting features such as Alcohol and 
Sulphates had an effect on the precieved quality of the red wine.  I was able to
find a correlation between quality and Alcohol and Sulphates. I was surprised to
find that Alcohol had a strong correlation with quality. I would have thought 
the opposite. 
One of the problems that I found with this dataset set that 
quality was more of a matter of opinion. As shown by some of the outliers, 
I think that a persons preference has an infulence on the quality when it came 
to certain features. For example, someone who likes sweeter wines is going to 
give a higher quality rating to wines with less citric acid. In the future 
I would try to come up with a more solid definition of quality. Maybe find 
quality ratings on a sub-category of red wine. Red wine seems to be a very broad
category of wine.


